Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-10-07 talk/pepm2011/paper.tex
some improvements

\begin{minipage}[b]{7 cm}
\emph{Object Domains:}
$$\setlength\arraycolsep{0.1em}
 \begin{array}{rcll}
    u,v,w & \in & V & \mathrm{\ variables\ in\ trace}\\
    u^*,v^*,w^* & \in & V^* & \mathrm{\ variables\ in\ optimized\ trace}\\
    T & \in & \mathfrak{T} & \mathrm{\ runtime\ types}\\
    F & \in & \left\{ L,R\right\} & \mathrm{\ fields\ of\ objects}\\
 \end{array}
$$
\end{minipage}
\begin{minipage}[b]{5 cm}
\emph{Semantic Values:}
$$\setlength\arraycolsep{0.1em}
 \begin{array}{rcll}
    E & \in & V\rightharpoonup V^* & \mathrm{\ Environment}\\
    S & \in & V^*\rightharpoonup\mathfrak{T}\times(V^*\cup \{ \mathrm{null} \})\times (V^*\cup \{ \mathrm{null}\} ) & \mathrm{\ Heap}\\
    \\
    \\
 \end{array}
$$
\end{minipage}
\end{center}
\label{fig:optimization}
\caption{Optimization Rules}
\end{figure*}

@@

In this section we described how simple escape analysis within the scope of one
loop works. This optimizations reduces the allocation of many intermediate data
structures that become garbage quickly in an interpreter. It also removes a lot
of the type dispatching overhead. In the next section, we will explain how this
optimization can be improved further.

@@

\section{Escape Analysis Across Loop Boundaries}
\label{sec:crossloop}

In the last section we described how escape analysis can be used to remove
many of the allocations of short-lived objects and many of the type dispatches
that are present in a non-optimized trace. In this section we will improve the
optimization to also handle more cases.

@@

Let's look at the final trace obtained in the last section for the example loop.
The final trace was much better than the original one, because many allocations
were removed from it. However, it also still contained allocations:

@@

Now the lifetime of the remaining allocations no longer crosses the jump, and
we can run our escape analysis a second time, to get the trace in
Figure~\ref{fig:step3}.

@@

% section Escape Analysis Across Loop Boundaries (end)

>>>>

\begin{minipage}[b]{7 cm}
\emph{Object Domains:}
$$\setlength\arraycolsep{0.1em}
 \begin{array}{rcll}
    u,v,w & \in & V & \mathrm{\ variables\ in\ trace}\\
    u^*,v^*,w^* & \in & V^* & \mathrm{\ variables\ in\ optimized\ trace}\\
    T & \in & \mathfrak{T} & \mathrm{\ runtime\ types}\\
    F & \in & \left\{ L,R\right\} & \mathrm{\ fields\ of\ objects}\\
 \end{array}
$$
\end{minipage}
\begin{minipage}[b]{5 cm}
\emph{Semantic Values:}
$$\setlength\arraycolsep{0.1em}
 \begin{array}{rcll}
    E & \in & V\rightharpoonup V^* & \mathrm{\ Environment}\\
    S & \in & V^*\rightharpoonup\mathfrak{T}\times(V^*\cup \{ \mathrm{null} \})\times (V^*\cup \{ \mathrm{null}\} ) & \mathrm{\ Static\ Heap}\\
    \\
    \\
 \end{array}
$$
\end{minipage}
\end{center}
\label{fig:optimization}
\caption{Optimization Rules}
\end{figure*}

@@

@@

\section{Allocation Removal Across Loop Boundaries}
\label{sec:crossloop}

In the last sections we described how partial evaluation can be used to remove
many of the allocations of short-lived objects and many of the type dispatches
that are present in a non-optimized trace. In this section we will improve the
optimization to also handle more cases.

@@

Let's look at the final trace obtained in Section~\ref{sec:statics} for the
example loop. The final trace (Figure~\ref{fig:step1}) was much better than the
original one, because many allocations were removed from it. However, it also
still contained allocations.

@@

Now the lifetime of the remaining allocations no longer crosses the jump, and
we can run our partial evaluation a second time, to get the trace in
Figure~\ref{fig:step3}.

@@

% section Allocation Removal Across Loop Boundaries (end)
